Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2019-10-01 manuscript_pretty.Rmd
editing

The key finding was that the congruency effects for the frequency unbiased items were significantly larger when the low-conflict context was made task-relevant as compared to when the high-conflict context was made task-relevant. This result is consistent with prior CSPC effects and, like the previous work, suggests that the context cues triggered rapid adjustments to attentional control [@crump_flexibility_2009]. However, unlike prior studies, we were able to experimentally manipulate the CSPC effect across blocks of trials without changing any of the physical properties of the stimuli. This novel finding demonstrates that participants were able to learn and apply one context-attention association in the first phase, and subsequently ignore that association to learn a new association in the second phase.

@@

Another key result of this study concerned the different stimuli used as context cues. Across the three conditions, we varied the type of context image and dimensions. We manipulated the type of image presented, including both objects (identity and color dimensions) and faces (gender and racial dimensions). We also manipulated whether a single set of three repeating images were presented (object and social) or a set of non-repeating images were presented (social/non-repeating). Across all three conditions, we found no evidence that using different stimuli had an influence on the size or direction of the CSPC demonstrating generalizability of this phenomenon. Furthermore, CSPC effects were present even when using non-repeating images which suggests that context-dependency did not rely on image-specific associations but higher-order, learned categorical information.

>>>>

The key finding, in Experiment 1, was that congruency effects for the frequency unbiased items were significantly larger when the low-conflict context was made task-relevant as compared to when the high-conflict context was made task-relevant. This result is consistent with prior CSPC effects and, like the previous work, suggests that the context cues triggered rapid adjustments to attentional control [@crump_flexibility_2009]. However, unlike prior studies, we were able to experimentally manipulate the CSPC effect across blocks of trials without changing any of the physical properties of the stimuli. This novel finding demonstrates that participants were able to learn and apply one context-attention association in the first phase, and subsequently ignore that association to learn a new association in the second phase.

@@

In Experiment 2, we found that participants did not spontaneously learn the associations between proportion congruency and context-cues without the secondary counting task. This replicates previous work () and demonstrates further that task-relevance is necessary for establishing associations between contexts and attentional priorities. Additionally, in Experiment 3 explicit instructions to make use of the cues also failed to produce context-specific effects. This result also tracks well with previous work that found instruction manipulations could successfully produce list-wide proportion congruency effects, but not item-specific () or context-specific () effects.

Another key result of this study concerned the different stimuli used as context cues in Experiment 1. Across the three conditions, we varied the type of context image and dimensions. We manipulated the type of image presented, including both objects (identity and color dimensions) and faces (gender and racial dimensions). We also manipulated whether a single set of three repeating images were presented (object and social) or a set of non-repeating images were presented (social/non-repeating). Across all three conditions, we found no evidence that using different stimuli had an influence on the size or direction of the CSPC demonstrating generalizability of this phenomenon. Furthermore, CSPC effects were present even when using non-repeating images which suggests that context-dependency did not rely on image-specific associations but higher-order, learned categorical information. We should also note an important limitation is that the experimental design, while moderately powered to detect the presence versus absence of the CSPC effect between groups, it was likely under-powered to detect small changes in the CSPC effect (see the Replicability section above).
